---
title: "CI-project-chemical"
output: html_document
date: '2022-12-08'
---

## Install packages & library
```{r}
#install.packages("remotes")
#remotes::install_github("ygeunkim/propensityml")
#install.packages("dplyr")
#install.packages("sas7bdat")

library(propensityml)
library(dplyr)
library(sas7bdat)

```
# Chemical Dataset
## Data reading
```{r}
df<-read.table('/Users/kyungseonlee/Desktop/causal/poisox.txt')
colnames(df)=c('age','sex','prior','poisox','subseq','mortal')
df%>% head()

df3=df
df3$blood_diff=df3$subseq-df3$prior
df3=df3[c(1,2,4,6,7)]
df3 %>% head()


df0=df[c(1,2,4,6)]
df0 %>% head()

```
## Propensity score estimation- logistic regression, random Forest, CART

## 1. Logistic Regression
### 1-a. LR-ps estimation
```{r}
log_reg_ps=glm(poisox~ age+sex+blood_diff, family = "binomial", data=df3)
logit_e_hat=predict(log_reg_ps)
print(logit_e_hat %>% head())

#exp(logit_e_hat)
e_hat=exp(logit_e_hat)/(1+exp(logit_e_hat))
print(e_hat %>% head())

df_lo=df3

df_lo[,"ps"]=e_hat
df_lo %>% head()

```
### 1-b. LR-weighting
```{r}


zi=df_lo$poisox
yi=df_lo$mortal
e=df_lo$ps

df_lo["ipw_wt"]= zi/e-(1-zi)/(1-e)
df_lo %>% head()

ATE_ipw_log=mean(zi*yi/e)-mean((1-zi)*yi/(1-e))
ATE_ipw_log

ATE_sipw_log=sum(zi*yi/e)/sum(zi/e)-sum((1-zi)*yi/(1-e))/sum((1-zi)/(1-e))
ATE_sipw_log

```
### 1-c. LR-Evaluate
```{r}
cov_balance_lo=data.frame(rep(0),row.names = "Logistic Regression")

df3 %>% head()
for(i in colnames(df3)){
  if(i!="poisox" & i!="mortal"){
#    print(df1[i])
    t_weighted_mean=mean((df3[i]*df_lo$ipw_wt)[df3$poisox==1,])
    c_weighted_mean=mean((df3[i]*df_lo$ipw_wt)[df3$poisox==0,])
    weighted_mean_diff=abs(t_weighted_mean-c_weighted_mean)
    asam=weighted_mean_diff/sd((df3[i]*df_lo$ipw_wt)[df3$poisox==1,])
    cov_balance_lo[i]=asam
  }
}
cov_balance_lo=cov_balance_lo[,-1]
cov_balance_lo
cov_balance_lo$ASAM=apply(cov_balance_lo,1,mean)
cov_balance_lo["ASAM"]
```
## 2. Random Forest
### 2-a. RF-ps estimation
```{r}

#install.packages("randomForest")
library(randomForest)

?randomForest

rf=randomForest(poisox~ age+sex+blood_diff, data=df3)
rf_ps=predict(rf)
print(rf_ps %>% head())

df_rf=df3

df_rf[,"ps"]=rf_ps
df_rf %>% head()

```

### 2-b. RF-weighting
```{r}

zi=df_rf$poisox
yi=df_rf$mortal
e=df_rf$ps

df_rf["ipw_wt"]= zi/e-(1-zi)/(1-e)
df_rf %>% head()

ATE_ipw_rf=mean(zi*yi/e)-mean((1-zi)*yi/(1-e))
ATE_ipw_rf

ATE_sipw_rf=sum(zi*yi/e)/sum(zi/e)-sum((1-zi)*yi/(1-e))/sum((1-zi)/(1-e))
ATE_sipw_rf


```
### 2-c. RF-Evaluation
```{r}
cov_balance_rf=data.frame(rep(0),row.names = "Random Forest")

df3 %>% head()
for(i in colnames(df3)){
  if(i!="poisox" & i!="mortal"){
#    print(df1[i])
    t_weighted_mean=mean((df3[i]*df_rf$ipw_wt)[df3$poisox==1,])
    c_weighted_mean=mean((df3[i]*df_rf$ipw_wt)[df3$poisox==0,])
    weighted_mean_diff=abs(t_weighted_mean-c_weighted_mean)
    asam=weighted_mean_diff/sd((df3[i]*df_rf$ipw_wt)[df3$poisox==1,])
    cov_balance_rf[i]=asam
  }
}
cov_balance_rf=cov_balance_rf[,-1]
cov_balance_rf
cov_balance_rf$ASAM=apply(cov_balance_rf,1,mean)
cov_balance_rf["ASAM"]
```

## 3. CART
### 3-a. CART-ps estimation

```{r}
library(rpart)

#control = rpart.control(minbucket = 2)
df1_c=df3
df1_c %>% head(5)

cart2=rpart(poisox~ age+sex+blood_diff+blood_diff, data=df1_c ,method='poisson',control = rpart.control(maxdepth = 1))
#summary(cart2)
cart2$cptable

cart_ps2=predict(cart2)
cart_ps2 %>% head()
df1_c[,"ps2"]=cart_ps2

df1_c %>% head()

sum(df1_c$ps2==0|df1_c$ps2==1)
```

```{r}
#install.packages("rpart.plot")

library(rpart.plot)

prp(cart2, type=4, extra=2, digits=3) 
```
### 3-b. CART-weighting
```{r}
zi=df1_c$poisox
yi=df1_c$mortal
e=df1_c$ps2


df1_c["ipw_wt"]= zi/e-(1-zi)/(1-e)
df1_c %>% head()

ATE_ipw_cart=mean(zi*yi/e)-mean((1-zi)*yi/(1-e))
ATE_ipw_cart
```

```{r}

ATE_sipw_cart=sum(zi*yi/e)/sum(zi/e)-sum((1-zi)*yi/(1-e))/sum((1-zi)/(1-e))
ATE_sipw_cart
```
### 3-c. CART-evaluation
```{r}
cov_balance_c=data.frame(rep(0),row.names = "CART")

df3%>% head()
for(i in colnames(df3)){
  if(i!="poisox" & i!="mortal"){
#    print(df1[i])
    t_weighted_mean=mean((df3[i]*df1_c$ipw_wt)[df3$poisox==1,])
    c_weighted_mean=mean((df3[i]*df1_c$ipw_wt)[df3$poisox==0,])
    weighted_mean_diff=abs(t_weighted_mean-c_weighted_mean)
    asam=weighted_mean_diff/sd((df3[i]*df1_c$ipw_wt)[df3$poisox==1,])
    cov_balance_c[i]=asam
  }
}
cov_balance_c=cov_balance_c[,-1]
cov_balance_c
cov_balance_c$ASAM=apply(cov_balance_c,1,mean)
cov_balance_c["ASAM"]
```

## Total ATE table
```{r}
ATE_table= rbind(ATE_ipw_log,
      ATE_sipw_log,
      ATE_ipw_rf,
      ATE_sipw_rf,
      ATE_ipw_cart, 
      ATE_sipw_cart)

colnames(ATE_table)="ATE table in Chemical Dataset"
knitr :: kable(ATE_table,"simple")
```

## Evaluation visualization - Chemical dataset
### a. ASAM table
```{r}
ASAM_table1=rbind(cov_balance_lo["ASAM"] ,cov_balance_rf["ASAM"] ,cov_balance_c["ASAM"])
colnames(ASAM_table1)="ASAM in Chemical Dataset"
knitr::kable(ASAM_table1,"simple")
```

### b. ASAM box plot
```{r}
a=cbind(t(cov_balance_lo) ,t(cov_balance_rf) ,t(cov_balance_c))
colnames(a)=c("LR","RF","CART")
boxplot(a, main="ASAM in the Chemical Dataset",horizontal = TRUE)
```

### c. ps distribution 
```{r}
b=cbind(df_lo["ps"],df_rf["ps"],df1_c["ps2"])
colnames(b)=c("LR","RF","CART")
boxplot(b,horizontal = TRUE,main="Propensity score distribution in the Chemical Dataset")
```

### d. weight distribution
```{r}
par(mfcol=c(1,3))
boxplot(df_lo[(df3$poisox==0),"ipw_wt"],main="LR in Chemical")
boxplot(df_rf[(df3$poisox==0),"ipw_wt"],main="RF in Chemical")
boxplot(df1_c[(df3$poisox==0),"ipw_wt"],main="CART in Chemical")

```

