---
title: "CI-project-lalonde"
output: html_document
date: '2022-12-15'
---
## Install packages & library
```{r}
#install.packages("remotes")
#remotes::install_github("ygeunkim/propensityml")
#install.packages("dplyr")
#install.packages("sas7bdat")
#install.packages("MatchIt")
#install.packages("survey")
#install.packages("tableone")

library(propensityml)
library(dplyr)
library(sas7bdat)
library(MatchIt)

library(ggplot2) 
library(moonBook) 
library(survey)
library(tableone)
```
# LaLonde Dataset
### 직업훈련프로그램(treat)를 들은 사람의 수입(outcome='re78')이 더 높은가?
## Data reading
```{r}
df<-lalonde
colnames(df)[9]<-"outcome"
df %>% head()
dim(df)

```
## Propensity score estimation- logistic regression, random Forest, CART

## 1. Logistic Regression
### 1-a. LR-ps estimation
```{r}
df_lo=df

log_reg_ps=glm(treat~ .-outcome, family = "binomial", data=df)
df_lo$ps=log_reg_ps$fitted.values
df_lo %>% head()

```
### 1-b. LR-weighting
```{r}
zi=df_lo$treat
yi=df_lo$outcome
e=df_lo$ps

df_lo$w<-ifelse(df_lo$treat==1,1/df_lo$ps,1/(1-df_lo$ps))
df_lo %>% head()
```


```{r}
ATE_ipw_log=mean(zi*yi/e)-mean((1-zi)*yi/(1-e))
ATE_ipw_log

ATE_sipw_log=sum(zi*yi/e)/sum(zi/e)-sum((1-zi)*yi/(1-e))/sum((1-zi)/(1-e))
ATE_sipw_log

```
### 1-c. LR-Evaluate
### 가중치 부여 전후 propensity score 분포 차이 비교.
```{r}
par(mfcol=c(1,2))
ggplot(data=df_lo, aes(x=ps))+geom_density(aes(fill=factor(treat)),alpha=0.3)

ggplot(data=df_lo, aes(x=ps, weight=w))+geom_density(aes(fill=factor(treat)),alpha=0.3) 
```

```{r}
cov_balance_lo=data.frame(rep(0),row.names = "Logistic Regression")
summary(df[,4])

for(i in colnames(df)){
  if(i!="treat" & i!="outcome" &i!="race"){
#    print(df1[i])
    t_weighted_mean=mean((df[i]*df_lo$w)[df$treat==1,])
    c_weighted_mean=mean((df[i]*df_lo$w)[df$treat==0,])
    weighted_mean_diff=abs(t_weighted_mean-c_weighted_mean)
    asam=weighted_mean_diff/sd((df[i]*df_lo$w)[df$treat==1,])
    cov_balance_lo[i]=asam
  }
}
cov_balance_lo=cov_balance_lo[,-1]
cov_balance_lo
cov_balance_lo$ASAM=apply(cov_balance_lo,1,mean)
cov_balance_lo["ASAM"]
```
## 2. Random Forest
### 2-a. RF-ps estimation
```{r}

#install.packages("randomForest")
library(randomForest)
df_rf=df

?randomForest

rf=randomForest(treat~ .-outcome, data=df)
df_rf$ps=predict(rf)
df_rf %>% head()

```

### 2-b. RF-weighting
```{r}

zi=df_rf$treat
yi=df_rf$outcome
e=df_rf$ps

df_rf$w<-ifelse(df_rf$treat==1,1/df_rf$ps,1/(1-df_rf$ps))
df_rf %>% head()

ATE_ipw_rf=mean(zi*yi/e)-mean((1-zi)*yi/(1-e))
ATE_ipw_rf

ATE_sipw_rf=sum(zi*yi/e)/sum(zi/e)-sum((1-zi)*yi/(1-e))/sum((1-zi)/(1-e))
ATE_sipw_rf


```
### 2-c. RF-Evaluation
### 가중치 부여 전후 propensity score 분포 차이 비교.
```{r}
par(mfcol=c(1,2))
ggplot(data=df_rf, aes(x=ps))+geom_density(aes(fill=factor(treat)),alpha=0.3)

ggplot(data=df_rf, aes(x=ps, weight=w))+geom_density(aes(fill=factor(treat)),alpha=0.3) 
```

```{r}
cov_balance_rf=data.frame(rep(0),row.names = "Random Forest")

df %>% head()
for(i in colnames(df)){
  if(i!="treat" & i!="outcome"&i!="race"){
#    print(df1[i])
    t_weighted_mean=mean((df[i]*df_rf$w)[df$treat==1,])
    c_weighted_mean=mean((df[i]*df_rf$w)[df$treat==0,])
    weighted_mean_diff=abs(t_weighted_mean-c_weighted_mean)
    asam=weighted_mean_diff/sd((df[i]*df_rf$w)[df$treat==1,])
    cov_balance_rf[i]=asam
  }
}
cov_balance_rf=cov_balance_rf[,-1]
cov_balance_rf
cov_balance_rf$ASAM=apply(cov_balance_rf,1,mean)
cov_balance_rf["ASAM"]
```

## 3. CART
### 3-a. CART-ps estimation

```{r}
library(rpart)

#control = rpart.control(minbucket = 2)
df_c<-df

cart=rpart(treat~ .-outcome, data=df_c ,method='poisson',control = rpart.control(minsplit=40))
#summary(cart2)
cart$cptable
df_c$ps<-predict(cart)

df_c %>% head()

sum(df_c$ps==0|df_c$ps==1)
```

```{r}
#install.packages("rpart.plot")

library(rpart.plot)

prp(cart, type=4, extra=2, digits=3) 
```
### 3-b. CART-weighting
```{r}
zi=df_c$treat
yi=df_c$outcome
e=df_c$ps


df_c$w<-ifelse(df_c$treat==1,1/df_c$ps,1/(1-df_c$ps))
df_c %>% head()

ATE_ipw_cart=mean(zi*yi/e)-mean((1-zi)*yi/(1-e))
ATE_ipw_cart
```

```{r}

ATE_sipw_cart=sum(zi*yi/e)/sum(zi/e)-sum((1-zi)*yi/(1-e))/sum((1-zi)/(1-e))
ATE_sipw_cart
```
### 3-c. CART-evaluation
### 가중치 부여 전후 propensity score 분포 차이 비교.
```{r}
par(mfcol=c(1,2))
ggplot(data=df_c, aes(x=ps))+geom_density(aes(fill=factor(treat)),alpha=0.3)

ggplot(data=df_c, aes(x=ps, weight=w))+geom_density(aes(fill=factor(treat)),alpha=0.3) 
```

```{r}
cov_balance_c=data.frame(rep(0),row.names = "CART")

df%>% head()
for(i in colnames(df)){
  if(i!="treat" & i!="outcome"&i!="race"){
#    print(df1[i])
    t_weighted_mean=mean((df[i]*df_c$w)[df$treat==1,])
    c_weighted_mean=mean((df[i]*df_c$w)[df$treat==0,])
    weighted_mean_diff=abs(t_weighted_mean-c_weighted_mean)
    asam=weighted_mean_diff/sd((df[i]*df_c$w)[df$treat==1,])
    cov_balance_c[i]=asam
  }
}
cov_balance_c=cov_balance_c[,-1]
cov_balance_c
cov_balance_c$ASAM=apply(cov_balance_c,1,mean)
cov_balance_c["ASAM"]
```

## Total ATE table
```{r}
ATE_table= rbind(ATE_ipw_log,
      ATE_sipw_log,
      ATE_ipw_rf,
      ATE_sipw_rf,
      ATE_ipw_cart, 
      ATE_sipw_cart)

colnames(ATE_table)="ATE table in LaLonde Dataset"
knitr :: kable(ATE_table,"simple")
```

## Evaluation visualization - LaLonde dataset
### a. ASAM table
```{r}
ASAM_table1=rbind(cov_balance_lo["ASAM"] ,cov_balance_rf["ASAM"] ,cov_balance_c["ASAM"])
ASAM_table1
colnames(ASAM_table1)="ASAM in LaLonde Dataset"
knitr::kable(ASAM_table1,"simple")
```

### b. ASAM box plot
```{r}
a=cbind(t(cov_balance_lo) ,t(cov_balance_rf) ,t(cov_balance_c))
colnames(a)=c("LR","RF","CART")
boxplot(a, main="ASAM in the LaLonde Dataset",horizontal = TRUE)
```

### c. ps distribution 


```{r}
b=cbind(df_lo["ps"],df_rf["ps"],df_c["ps"])
colnames(b)=c("LR","RF","CART")
boxplot(b,horizontal = TRUE,main="Propensity score distribution in the LaLonde Dataset")
```

### d. weight distribution
```{r}
par(mfcol=c(1,3))
boxplot(df_lo[(df$treat==0),"w"],main="LR in LaLonde")
boxplot(df_rf[(df$treat==0),"w"],main="RF in LaLonde")
boxplot(df_c[(df$treat==0),"w"],main="CART in LaLonde")

```

