---
title: "Causal_project"
output: html_document
date: '2022-11-15'
---
## Install packages & library
```{r}
#install.packages("remotes")
#remotes::install_github("ygeunkim/propensityml")
#install.packages("dplyr")
#install.packages("sas7bdat")

library(propensityml)
library(dplyr)
library(sas7bdat)

```


## Korea National Health and Nutrition Examination Survey Dataset
## 1. Data reading
```{r}

raw_data <- read.sas7bdat("/Users/kyungseonlee/snu-causal/main/input/rawdata/HN20_ALL.sas7bdat")
dim(raw_data) %>% head()
#raw_data %>% head(1)

```

# 2. Data preprocessing
```{r}
column=c('sex','age','BD1','HE_ast','HE_alt',"HE_BMI",'DK8_dg','DK9_dg')
#column2=c('BD1','DC2_dg','DK8_dg','DK9_dg','DK4_dg')

df=raw_data[,column]
df %>% head(7)

sum(df[,c(8)])
```


## Missing value deletion
```{r}
for(i in 3:ncol(df)){
  if(i==4|i==5|i==6){
    df<- df[!( is.na(df[,i])), ]
  }else{
    df<- df[!(df[,i] == 8 |df[,i] == 9 | is.na(df[,i])), ]
  }
}
for (i in 1:nrow(df)){
  if (df[i,7]+df[i,8]==0){df[i,9]=0}else{df[i,9]=1}
}
df %>% head(20)
sum(is.na(df))
dim(df)
```


```{r}
# First topic: Hepatitis causal inference
df1=df[,c(-7,-8)]
colnames(df1)[3:ncol(df1)]=c("treat","ast","alt","bmi","outcome")

df1 %>% head(7)

```
## Diabetes & Drinking

```{r}
df1$treat<-df1$treat-1
df1$sex<-df1$sex-1

df1 %>% head(7)

table(df1$treat,df1$outcome)

# (df1$DC8_dg==0)
# sum(is.na(df1$treat))
# sum(is.na(df1$outcome))
# dim(df1)
# which(df1$treat==0) %>% head()
```

## Propensity score estimation- logistic regression, random Forest, CART

## 1. Logistic Regression
### 1-a. LR-ps estimation
```{r}
#logsitic regression
df1_lo=df1
df1_lo %>% head(10)

log_reg_ps=glm(treat~ .-outcome, family = "binomial", data=df1)
logit_e_hat=predict(log_reg_ps)
print(logit_e_hat %>% head())

lo_ps=exp(logit_e_hat)/(1+exp(logit_e_hat))
print(lo_ps %>% head())

df1_lo[,"ps"]=lo_ps
# 
# for (i in 1:nrow(df1_lo) ){
#   if (df1_lo$treat[i]==1){
#     df1_lo[i,"ps"]=lo_ps[i]
#   }
#   else{
#     df1_lo[i,"ps"]=lo_ps[i]
#   }
# }
df1_lo %>% head()
```

### 1-b. LR-weighting

```{r}
zi=df1_lo$treat
yi=df1_lo$outcome
e=df1_lo$ps

df1_lo["ipw_wt"]= zi/e-(1-zi)/(1-e)
df1_lo %>% head()

ATE_ipw_log=mean(zi*yi/e)-mean((1-zi)*yi/(1-e))
ATE_ipw_log

ATE_sipw_log=sum(zi*yi/e)/sum(zi/e)-sum((1-zi)*yi/(1-e))/sum((1-zi)/(1-e))
ATE_sipw_log

```
### 1-c. LR-Evaluate

```{r}
cov_balance_lo=data.frame(rep(0),row.names = "logistic regression")

for(i in colnames(df1)){
  if(i!="treat" & i!="outcome"){
#    print(df1[i])
    t_weighted_mean=mean((df1[i]*df1_lo$ipw_wt)[df1$treat==1,])
    c_weighted_mean=mean((df1[i]*df1_lo$ipw_wt)[df1$treat==0,])
    weighted_mean_diff=abs(t_weighted_mean-c_weighted_mean)
    asam=weighted_mean_diff/sd((df1[i]*df1_lo$ipw_wt)[df1$treat==1,])
    cov_balance_lo[i]=asam
  }
}
cov_balance_lo=cov_balance_lo[,-1]
cov_balance_lo$ASAM=apply(cov_balance_lo,1,mean)
cov_balance_lo["ASAM"]

```

## 2. Random Forest
### 2-a. RF-ps estimation
```{r}
library(randomForest)
df1_rf=df1

rf=randomForest(treat~ .-outcome, data=df1)
rf_ps=predict(rf)
print(rf_ps %>% head())

df1_rf[,"ps"]=rf_ps

# for (i in 1:nrow(df1_rf) ){
#   if (df1_rf$treat[i]==1){
#     df1_rf[i,"ps"]=rf_ps[i]
#   }
#   else{
#     df1_rf[i,"ps"]=rf_ps[i]
#   }
# }
df1_rf %>% head()
```

### 2-b. RF-weighting

```{r}
zi=df1_rf$treat
yi=df1_rf$outcome
e=df1_rf$ps

df1_rf["ipw_wt"]= zi/e-(1-zi)/(1-e)
df1_rf %>% head()

#ATE estimation
ATE_ipw_rf=mean(zi*yi/e)-mean((1-zi)*yi/(1-e))
ATE_ipw_rf

ATE_sipw_rf=sum(zi*yi/e)/sum(zi/e)-sum((1-zi)*yi/(1-e))/sum((1-zi)/(1-e))
ATE_sipw_rf
```

### 2-c. RF-Evaluation
```{r}

cov_balance_rf=data.frame(rep(0),row.names = "Random Forest")

for(i in colnames(df1)){
  if(i!="treat" & i!="outcome"){
#    print(df1[i])
    t_weighted_mean=mean((df1[i]*df1_rf$ipw_wt)[df1$treat==1,])
    c_weighted_mean=mean((df1[i]*df1_rf$ipw_wt)[df1$treat==0,])
    weighted_mean_diff=abs(t_weighted_mean-c_weighted_mean)
    asam=weighted_mean_diff/sd((df1[i]*df1_rf$ipw_wt)[df1$treat==1,])
    cov_balance_rf[i]=asam
  }
}
cov_balance_rf=cov_balance_rf[,-1]
cov_balance_rf
cov_balance_rf$ASAM=apply(cov_balance_rf,1,mean)
cov_balance_rf["ASAM"]
```
## 3. CART
### 3-a. CART-ps estimation
```{r}
#install.packages("rpart")
library(rpart)

# cart=rpart(poisox~ .-outcome, data=df10 ,method='poisson')
# summary(cart)
# 
# df10_c=df10
# cart_ps=predict(cart)
# df10_c["propensity score"]=cart_ps
# df10_c %>% head()
# 
# df11_c=df11

df1_c=df1
df1_c %>% head(5)

cart=rpart(treat~ .-outcome, data=df1_c ,method='poisson',control = rpart.control(maxdepth = 5))
#summary(cart2)
cart$cptable

cart_ps=predict(cart)
cart_ps %>% head()
df1_c[,"ps"]=cart_ps

df1_c %>% head()

sum(df1_c$ps==0|df1_c$ps==1)

#df1_c[65:70,]
```


```{r}
#install.packages("rpart.plot")

library(rpart.plot)

prp(cart, type=4, extra=2, digits=3) 
```
### 3-b. CART-weighting

```{r}
#add weighting- CART
zi=df1_c$treat
yi=df1_c$outcome
e=df1_c$ps

df1_c["ipw_wt"]= zi/e-(1-zi)/(1-e)
df1_c %>% head()

#ATE_ipw
ATE_ipw_cart=mean(zi*yi/e)-mean((1-zi)*yi/(1-e))
ATE_ipw_cart

#ATE_sipw
ATE_sipw_cart=sum(zi*yi/e)/sum(zi/e)-sum((1-zi)*yi/(1-e))/sum((1-zi)/(1-e))
ATE_sipw_cart
```

### 3-c. CART-evaluation
```{r}

cov_balance_cart=data.frame(rep(0),row.names = "CART")

for(i in colnames(df1)){
  if(i!="treat" & i!="outcome"){
#    print(df1[i])
    t_weighted_mean=mean((df1[i]*df1_c$ipw_wt)[df1$treat==1,])
    c_weighted_mean=mean((df1[i]*df1_c$ipw_wt)[df1$treat==0,])
    weighted_mean_diff=abs(t_weighted_mean-c_weighted_mean)
    asam=weighted_mean_diff/sd((df1[i]*df1_c$ipw_wt)[df1$treat==1,])
    cov_balance_cart[i]=asam
  }
}
cov_balance_cart=cov_balance_cart[,-1]
cov_balance_cart$ASAM=apply(cov_balance_cart,1,mean)
cov_balance_cart["ASAM"]
```
## Total ATE table

```{r}
ATE_table= rbind(ATE_ipw_log,
      ATE_sipw_log,
      ATE_ipw_rf,
      ATE_sipw_rf,
      ATE_ipw_cart, 
      ATE_sipw_cart)

colnames(ATE_table)="ATE table in KHN Dataset"

knitr :: kable(ATE_table,"simple")
```

```{r, results='asis'}

#install.packages("knitr")

```


## Evaluation visualization - Chemical dataset
### a. ASAM table
```{r}
ASAM_table2=rbind(cov_balance_lo["ASAM"] ,cov_balance_rf["ASAM"] ,cov_balance_cart["ASAM"])
colnames(ASAM_table2)="ASAM in KHN Dataset"
knitr::kable(ASAM_table2,"simple")
```

### b. ASAM box plot
```{r}
a=cbind(t(cov_balance_lo) ,t(cov_balance_rf) ,t(cov_balance_cart))
colnames(a)=c("LR","RF","CART")
boxplot(a, main="ASAM in the KHN Dataset",horizontal = TRUE)
```

### c. ps distribution
```{r}
b=cbind(df1_lo["ps"],df1_rf["ps"],df1_c["ps"])
colnames(b)=c("LR","RF","CART")
boxplot(b,horizontal = TRUE,main="Propensity score distribution in the KHN Dataset")
```

### d. weight distribution
```{r}
par(mfcol=c(1,3))
boxplot(df1_lo[(df1_lo$treat==0),"ipw_wt"],main="LR in KHN")
boxplot(df1_rf[(df1_lo$treat==0),"ipw_wt"],main="RF in KHN")
boxplot(df1_c[(df1_lo$treat==0),"ipw_wt"],main="CART in KHN")
```

```{r}
sim=data.frame(c(0.094,0.075,0.143))
colnames(sim)="ASAM in simulated dataset"
rownames(sim)=c("logistic regression","Random Forest","CART")
knitr :: kable(sim,"simple")
```


