---
title: "Causal_project"
output: html_document
date: '2022-11-15'
---
  ## Install packages & library
```{r}
#install.packages("remotes")
#remotes::install_github("ygeunkim/propensityml")
#install.packages("dplyr")
#install.packages("sas7bdat")

library(propensityml)
library(dplyr)
library(sas7bdat)

```
# 1. Chemical Dataset
## 1-1. Data reading
```{r}
df<-read.table('./poisox.txt')
colnames(df)=c('age','sex','prior','poisox','subseq','mortal')
df%>% head()

#df0 has integer values. : int값을 가진 sex, poisox
df0=df[c(1,2,4,6)]
df0 %>% head()

#df1 has factor values. : facotr 형태의 sex, poisox
df1=df0
df1[2:3]=lapply(df1[2:3],factor)
df1 %>% head()
```
## 1-2. Propensity score estimation- logistic regression, random Forest, CART
```{r}
# a. propensity score estimation-logistic regression
log_reg_ps=glm(poisox~ age+sex, family = "binomial", data=df)
logit_e_hat=predict(log_reg_ps)
print(logit_e_hat %>% head())

#exp(logit_e_hat)
e_hat=exp(logit_e_hat)/(1+exp(logit_e_hat))

e_hat %>% head()


df_glm <- df0
for (i in 1:nrow(df_glm) ){
  if (df_glm$poisox[i]==1){
    df_glm[i,"ps"]=e_hat[i]
  } #df_rf의 i번째 행, ps 열의 값이 rf_ps[i]로 지정된다.
  else{
    df_glm[i,"ps"]=e_hat[i]
  }
}
df_glm%>%head()


#Logistic Regression- weighting

zi=df_glm$poisox
yi=df_glm$mortal
e=df_glm$ps

df_glm["ipw_wt"]= zi/e-(1-zi)/(1-e)
df_glm %>% head()

ATE_ipw_glm=mean(zi*yi/e)-mean((1-zi)*yi/(1-e))

ATE_sipw_glm=sum(zi*yi/e)/sum(zi/e)-sum((1-zi)*yi/(1-e))/sum((1-zi)/(1-e))

df_glm%>%head()
```

```{r}
# a-1. propensity score estimation-logistic regression by package
log_reg_ps2=ps_glm(poisox~ age+sex, data=df)
e_hat2=estimate_ps(log_reg_ps2)
e_hat2 %>% head()

print(sum(e_hat!=e_hat2))
#직접 계산한것과 package로 한것의 차이가 없다는 것을 찾았습니다! 

```

```{r}
# b.  propensity score estimation- Random Forest
#install.packages("randomForest")
library(randomForest)
rf=randomForest(poisox~ age+sex, data=df0) ## factor 형태의 변수를 사용, classification
rf_ps=predict(rf)
print(rf_ps %>% head())

df_rf=df0

for (i in 1:nrow(df_rf) ){
  if (df_rf$poisox[i]==1){
    df_rf[i,"ps"]=rf_ps[i]
  } #df_rf의 i번째 행, ps 열의 값이 rf_ps[i]로 지정된다.
  else{
    df_rf[i,"ps"]=rf_ps[i]
  }
}
df_rf %>% head() #데이터셋에 ps 열을 추가한 data

#Random Forest- weighting

zi=df_rf$poisox
yi=df_rf$mortal
e=df_rf$ps

df_rf["ipw_wt"]= zi/e-(1-zi)/(1-e)
df_rf %>% head()

ATE_ipw_rf=mean(zi*yi/e)-mean((1-zi)*yi/(1-e))

ATE_sipw_rf=sum(zi*yi/e)/sum(zi/e)-sum((1-zi)*yi/(1-e))/sum((1-zi)/(1-e))

df_rf%>%head()

```

# b.  propensity score estimation- Random Forest by package


```{r}
#c. propensity score estimation- CART
#install.packages("rpart")
library(rpart)

# cart=rpart(poisox~ age+sex, data=df0 ,method='poisson')
# summary(cart)
# 
# df0_c=df0
# cart_ps=predict(cart)
# df0_c["propensity score"]=cart_ps
# df0_c %>% head()
# 
# df1_c=df1

df_cart=df1
cart=rpart(poisox~ age+sex, data=df_cart ,method='class')
summary(cart)

cart_ps=predict(cart)
cart_ps %>% head()
df_cart %>% head()
for (i in 1:nrow(df_cart) ){
  if (df_cart$poisox[i]=="1"){
    df_cart[i,"ps"]=cart_ps[i,"1"]
    df_cart[i,"trt"]=1}
  else{
    df_cart[i,"ps2"]=cart_ps[i,"0"]
    df_cart[i,"trt"]=0
  }
}
df_cart %>% head()
```


```{r}
#c. propensity score estimation- CART by package

cart2=ps_cart(poisox ~ age + sex, data = df1)
cart_ps2=estimate_ps(cart2) 
df1["propensity score"]=cart_ps2
df1 %>% head()

```
```{r}
# d. Add ipw_weight & calculate ATE_ipw
df_cart %>% head()
# for (i in 1:nrow(df1) ){
#   if (df1$poisox[i]=="1"){
#     df1[i,"propensity score2"]=cart_ps2[i,"0"]}
#   else{df1[i,"propensity score2"]=cart_ps2[i,"1"]}
# }
zi=df_cart$trt
yi=df_cart$mortal
e=df_cart$`propensity score`


df_cart["ipw_wt"]= zi/e-(1-zi)/(1-e)
df_cart %>% head()

```

```{r}
# d. calculate ATE_ipw and ATE_sipw

ATE_ipw_cart=mean(zi*yi/e)-mean((1-zi)*yi/(1-e))

ATE_sipw_cart=sum(zi*yi/e)/sum(zi/e)-sum((1-zi)*yi/(1-e))/sum((1-zi)/(1-e))
```

```{r}
#e. ATE_ipw and ATE_sipw of 'GLM, RF, CART'
method_123 <- c('glm', 'rf', 'cart')

ipw_sipw_method <- data.frame(method_123)
ipw_sipw_method$ATE_ipw <- c(ATE_ipw_glm, ATE_ipw_rf, ATE_ipw_cart)
ipw_sipw_method$ATE_sipw <- c(ATE_sipw_glm, ATE_sipw_rf, ATE_sipw_cart)

ipw_sipw_method
```


# 2. Korea National Health and Nutrition Examination Survey Dataset
## 2-1. Data reading
```{r}
raw_data <- read.sas7bdat("/Users/kyungseonlee/snu-causal/main/input/rawdata/HN20_ALL.sas7bdat")
dim(raw_data) %>% head()
#raw_data %>% head(1)

```

# 2-2. Data preprocessing
```{r}
column=c('BD1','DC2_dg','sex','age','DK8_dg','DK9_dg','DK4_dg')
column2=c('BD1','DC2_dg','DK8_dg','DK9_dg','DK4_dg')


df=raw_data[,column]
df %>% head(7)

df <- df[!(df$BD1 == 8 | df$BD1 ==9 | is.na(df$BD1)), ]
df <- df[!(df$DC2_dg == 8 | df$DC2_dg ==9 ), ]
df <- df[!(df$DK8_dg == 8 | df$DK8_dg ==9 ), ]
df <- df[!(df$DK9_dg == 8 | df$DK9_dg ==9 ), ]
df <- df[!(df$DK4_dg == 8 | df$DK4_dg ==9 ), ]

df %>% head(7)
```


```{r}
df$BD1<-df$BD1-1
df$sex<-df$sex-1

table(df$DC2_dg)
table(df$BD1)
sum(is.na(df$BD1))
sum(is.na(df$DC2_dg))
dim(df)
which(df$BD1==0) %>% head()
```

## ATE estimation - glm, random forest, cart

```{r}
#logsitic regression
df_lo=df

log_reg_ps=glm(BD1~ .-DC2_dg, family = "binomial", data=df)
logit_e_hat=predict(log_reg_ps)
print(logit_e_hat %>% head())

lo_ps=exp(logit_e_hat)/(1+exp(logit_e_hat))
print(lo_ps %>% head())


for (i in 1:nrow(df_lo) ){
  if (df_lo$BD1[i]==1){
    df_lo[i,"ps"]=lo_ps[i]
  }
  else{
    df_lo[i,"ps"]=lo_ps[i]
  }
}
df_lo %>% head()

#logistic regression - weighting

zi=df_lo$BD1
yi=df_lo$DC2_dg
e=df_lo$ps

df_lo["ipw_wt"]= zi/e-(1-zi)/(1-e)
df_lo %>% head()

ATE_ipw_log=mean(zi*yi/e)-mean((1-zi)*yi/(1-e))
ATE_ipw_log

ATE_sipw_log=sum(zi*yi/e)/sum(zi/e)-sum((1-zi)*yi/(1-e))/sum((1-zi)/(1-e))
ATE_sipw_log


```

```{r}
# random forest - propensity score estimation
df_rf=df

rf=randomForest(BD1~ .-DC2_dg, data=df)
rf_ps=predict(rf)
print(rf_ps %>% head())


for (i in 1:nrow(df_rf) ){
  if (df_rf$BD1[i]==1){
    df_rf[i,"ps"]=rf_ps[i]
  }
  else{
    df_rf[i,"ps"]=rf_ps[i]
  }
}
df_rf %>% head()

#Random Forest- weighting

zi=df_rf$BD1
yi=df_rf$DC2_dg
e=df_rf$ps

df_rf["ipw_wt"]= zi/e-(1-zi)/(1-e)
df_rf %>% head()

ATE_ipw_rf=mean(zi*yi/e)-mean((1-zi)*yi/(1-e))
ATE_ipw_rf

ATE_sipw_rf=sum(zi*yi/e)/sum(zi/e)-sum((1-zi)*yi/(1-e))/sum((1-zi)/(1-e))
ATE_sipw_rf
```

```{r}
#propensity score estimation- CART
#install.packages("rpart")
#library(rpart)

# cart=rpart(poisox~ age+sex, data=df0 ,method='poisson')
# summary(cart)
# 
# df0_c=df0
# cart_ps=predict(cart)
# df0_c["propensity score"]=cart_ps
# df0_c %>% head()
# 
# df1_c=df1

df_c=df
df_c["BD1_f"]=as.factor(df_c$BD1)

cart2=rpart(BD1~ .-DC2_dg-BD1_f, data=df_c ,method='class')
summary(cart2)

cart_ps2=predict(cart2)
cart_ps2 %>% head()
df_c %>% head()
df_c$BD1[95:100]

#add propensity score
for (i in 1:nrow(df_c) ){
  
  if (df_c$BD1[i]==1){
    df_c[i,"ps"]=cart_ps2[i,"1"]
  }
  else{
    df_c[i,"ps"]=cart_ps2[i,"0"]
  }
}

#df_c[65:70,]
```






```{r}
#add weighting- CART
zi=df_c$BD1
yi=df_c$DC2_dg
e=df_c$ps

df_c["ipw_wt"]= zi/e-(1-zi)/(1-e)
df_c %>% head()

#ATE_ipw
ATE_ipw_cart=mean(zi*yi/e)-mean((1-zi)*yi/(1-e))
ATE_ipw_cart

#ATE_sipw
ATE_sipw_cart=sum(zi*yi/e)/sum(zi/e)-sum((1-zi)*yi/(1-e))/sum((1-zi)/(1-e))
ATE_sipw_cart
# ATE_hat is very small..
```
```{r}
rbind(ATE_ipw_log,ATE_sipw_log,ATE_ipw_rf,ATE_sipw_rf,ATE_ipw_cart, ATE_sipw_cart)
```

```

### LaLonde dataset
```{r}
#
# haven::read_data("http://www.nber.org/~rdehejia/data/nsw_dw.dta")
#
# lalonde::nsw_dw
#
# data(LaLonde)
#
# print(data)
#
# y <- LaLonde$outcome
#
# trt <- LaLonde$treat
#
# x.varnames <- c("age", "educ", "black", "hisp", "white",
#                 "marr", "nodegr", "log.re75", "u75")
#
# # covariates
# data.x <- LaLonde[, x.varnames]
#
#
#
# # construct design matrix (with no intercept)
# x <- model.matrix(~ -1 + ., data = data.x)
#
# const.propens <- function(x, trt)
# {
#     mean.trt <- mean(trt == "Trt")
#     rep(mean.trt, length(trt))
# }
#
# subgrp_fit_w <- fit.subgroup(x = x, y = y, trt = trt,
#     loss = "logistic_loss_lasso",
#     propensity.func = const.propens,
#     cutpoint = 0,
#     type.measure = "auc",
#     nfolds = 10)
#
# summary(subgrp_fit_w)
```
