---
title: "Causal_project"
output: html_document
date: '2022-11-15'
---
## Install packages & library
```{r}
#install.packages("remotes")
#remotes::install_github("ygeunkim/propensityml")
#install.packages("dplyr")
#install.packages("sas7bdat")

library(propensityml)
library(dplyr)
library(sas7bdat)

```
# 1. Chemical Dataset
## 1-1. Data reading
```{r}
df<-read.table('/Users/kyungseonlee/Desktop/causal/poisox.txt')
colnames(df)=c('age','sex','prior','poisox','subseq','mortal')
df%>% head()

#df0 has integer values.
df0=df[c(1,2,4,6)]
df0 %>% head()

#df1 has factor values.
df1=df0
df1[2:3]=lapply(df1[2:3],factor)
df1 %>% head()
```
## 1-2. Propensity score estimation- logistic regression, random Forest, CART
```{r}
# a. propensity score estimation-logistic regression
log_reg_ps=glm(poisox~ age+sex, family = "binomial", data=df)
logit_e_hat=predict(log_reg_ps)
print(logit_e_hat %>% head())

#exp(logit_e_hat)
e_hat=exp(logit_e_hat)/(1+exp(logit_e_hat))
print(e_hat %>% head())
```

```{r}
# a-1. propensity score estimation-logistic regression by package

log_reg_ps2=ps_glm(poisox~ age+sex, data=df)
e_hat2=estimate_ps(log_reg_ps2)
e_hat2 %>% head()

print(sum(e_hat!=e_hat2))

```

```{r}
# b.  propensity score estimation- Random Forest
#install.packages("randomForest")
library(randomForest)
rf=randomForest(poisox~ age+sex, data=df0)
rf_ps=predict(rf)
print(rf_ps %>% head())

df_rf=df0

for (i in 1:nrow(df_rf) ){
  if (df_rf$poisox[i]==1){
    df_rf[i,"ps"]=rf_ps[i]
  }
  else{
    df_rf[i,"ps"]=rf_ps[i]
  }
}
df_rf %>% head()

#Random Forest- weighting

zi=df_rf$poisox
yi=df_rf$mortal
e=df_rf$ps

df_rf["ipw_wt"]= zi/e-(1-zi)/(1-e)
df_rf %>% head()

ATE_ipw=mean(zi*yi/e)-mean((1-zi)*yi/(1-e))
ATE_ipw

ATE_sipw=sum(zi*yi/e)/sum(zi/e)-sum((1-zi)*yi/(1-e))/sum((1-zi)/(1-e))
ATE_sipw





```

```{r}
# b.  propensity score estimation- Random Forest by package
```

```{r}
#c. propensity score estimation- CART
#install.packages("rpart")
library(rpart)

# cart=rpart(poisox~ age+sex, data=df0 ,method='poisson')
# summary(cart)
# 
# df0_c=df0
# cart_ps=predict(cart)
# df0_c["propensity score"]=cart_ps
# df0_c %>% head()
# 
# df1_c=df1

df1_c=df1
cart2=rpart(poisox~ age+sex, data=df1_c ,method='class')
summary(cart2)

cart_ps2=predict(cart2)
cart_ps2 %>% head()
df1_c %>% head()
for (i in 1:nrow(df1) ){
  if (df1$poisox[i]=="1"){
    df1_c[i,"ps2"]=cart_ps2[i,"1"]
    df1_c[i,"trt"]=1}
  else{
    df1_c[i,"ps2"]=cart_ps2[i,"0"]
    df1_c[i,"trt"]=0
  }
}
df1_c %>% head()

```


```{r}
#c. propensity score estimation- CART by package

cart2=ps_cart(poisox ~ age + sex, data = df1)
cart_ps2=estimate_ps(cart2)
df1["propensity score"]=cart_ps2
df1 %>% head()

```
```{r}
# d. Add ipw_weight & calculate ATE_ipw
df1_c %>% head()
# for (i in 1:nrow(df1) ){
#   if (df1$poisox[i]=="1"){
#     df1[i,"propensity score2"]=cart_ps2[i,"0"]}
#   else{df1[i,"propensity score2"]=cart_ps2[i,"1"]}
# }
zi=df1_c$trt
yi=df1_c$mortal
e=df1_c$ps2

df1_c["ipw_wt"]= zi/e-(1-zi)/(1-e)
df1_c %>% head()

ATE_ipw=mean(zi*yi/e)-mean((1-zi)*yi/(1-e))
ATE_ipw
```

```{r}
# d. calculate ATE_sipw

ATE_sipw=sum(zi*yi/e)/sum(zi/e)-sum((1-zi)*yi/(1-e))/sum((1-zi)/(1-e))
ATE_sipw

```


# 2. Korea National Health and Nutrition Examination Survey Dataset
## 2-1. Data reading
```{r}

raw_data <- read.sas7bdat("/Users/kyungseonlee/snu-causal/main/input/rawdata/HN20_ALL.sas7bdat")
dim(raw_data) %>% head()
#raw_data %>% head(1)

```

# 2-2. Data preprocessing
```{r}
column=c('BD1','sex','age','BD1_11','BD2_1','DE1_dg','DI1_dg')
column2=c('BD1','DC2_dg','DK8_dg','DK9_dg','DK4_dg')

df=raw_data[,column]
df %>% head(7)

for(i in 1:ncol(df)){
  if(i==4|i==5){
    df<- df[!(df[,i] == 9 | is.na(df[,i])), ]
    df[(df[,i]==8),i]=0
  }else{
    df<- df[!(df[,i] == 8 |df[,i] == 9 | is.na(df[,i])), ]
  }
  
}
df %>% head(20)
sum(is.na(df))

# First topic: Diabetes causal inference
df1=df[,-7]
colnames(df1)[1]="treat"
colnames(df1)[6]="outcome"

# Second topic: High blood pressure causal inference
df2=df[,-6]
colnames(df2)[1]="treat"
colnames(df2)[6]="outcome"

df1 %>% head(7)
df2 %>% head(7)
```
## Diabetes & Drinking

```{r}
df1$treat<-df1$treat-1
df1$sex<-df1$sex-1

df1 %>% head(7)

table(df1$treat,df1$outcome)

# (df1$DC8_dg==0)
# sum(is.na(df1$treat))
# sum(is.na(df1$outcome))
# dim(df1)
# which(df1$treat==0) %>% head()
```

## ATE estimation - glm, random forest, cart

```{r}
#logsitic regression
df1_lo=df1
df1_lo %>% head(10)

log_reg_ps=glm(treat~ sex+age, family = "binomial", data=df1)
logit_e_hat=predict(log_reg_ps)
print(logit_e_hat %>% head())

lo_ps=exp(logit_e_hat)/(1+exp(logit_e_hat))
print(lo_ps %>% head())


for (i in 1:nrow(df1_lo) ){
  if (df1_lo$treat[i]==1){
    df1_lo[i,"ps"]=lo_ps[i]
  }
  else{
    df1_lo[i,"ps"]=lo_ps[i]
  }
}
df1_lo %>% head()

#logistic regression - weighting

zi=df1_lo$treat
yi=df1_lo$outcome
e=df1_lo$ps

df1_lo["ipw_wt"]= zi/e-(1-zi)/(1-e)
df1_lo %>% head()

#ATE estimation
ATE_ipw_log=mean(zi*yi/e)-mean((1-zi)*yi/(1-e))
ATE_ipw_log

ATE_sipw_log=sum(zi*yi/e)/sum(zi/e)-sum((1-zi)*yi/(1-e))/sum((1-zi)/(1-e))
ATE_sipw_log

```
## Evaluate

```{r}
cov_balance_lo=data.frame(rep(0),row.names = "logistic regression")

for(i in colnames(df1)){
  if(i!="treat" & i!="outcome"){
#    print(df1[i])
    t_weighted_mean=mean((df1[i]*df1_lo$ipw_wt)[df1$treat==1,])
    c_weighted_mean=mean((df1[i]*df1_lo$ipw_wt)[df1$treat==0,])
    weighted_mean_diff=abs(t_weighted_mean-c_weighted_mean)
    asam=weighted_mean_diff/sd((df1[i]*df1_lo$ipw_wt)[df1$treat==1,])
    cov_balance_lo[i]=asam
  }
}
cov_balance_lo=cov_balance_lo[,-1]
cov_balance_lo$Average_ASAM=apply(cov_balance_lo,1,mean)
cov_balance_lo["Average_ASAM"]

```

```{r}
# random forest - propensity score estimation
df1_rf=df1

rf=randomForest(treat~ age+sex, data=df1)
rf_ps=predict(rf)
print(rf_ps %>% head())


for (i in 1:nrow(df1_rf) ){
  if (df1_rf$treat[i]==1){
    df1_rf[i,"ps"]=rf_ps[i]
  }
  else{
    df1_rf[i,"ps"]=rf_ps[i]
  }
}
df1_rf %>% head()

#Random Forest- weighting

zi=df1_rf$treat
yi=df1_rf$outcome
e=df1_rf$ps

df1_rf["ipw_wt"]= zi/e-(1-zi)/(1-e)
df1_rf %>% head()

#ATE estimation
ATE_ipw_rf=mean(zi*yi/e)-mean((1-zi)*yi/(1-e))
ATE_ipw_rf

ATE_sipw_rf=sum(zi*yi/e)/sum(zi/e)-sum((1-zi)*yi/(1-e))/sum((1-zi)/(1-e))
ATE_sipw_rf
```

## Evaluate
```{r}

cov_balance_rf=data.frame(rep(0),row.names = "Random Forest")

for(i in colnames(df1)){
  if(i!="treat" & i!="outcome"){
#    print(df1[i])
    t_weighted_mean=mean((df1[i]*df1_rf$ipw_wt)[df1$treat==1,])
    c_weighted_mean=mean((df1[i]*df1_rf$ipw_wt)[df1$treat==0,])
    weighted_mean_diff=abs(t_weighted_mean-c_weighted_mean)
    asam=weighted_mean_diff/sd((df1[i]*df1_rf$ipw_wt)[df1$treat==1,])
    cov_balance_rf[i]=asam
  }
}
cov_balance_rf=cov_balance_rf[,-1]
cov_balance_rf$Average_ASAM=apply(cov_balance_rf,1,mean)
cov_balance_rf["Average_ASAM"]
```

```{r}
#propensity score estimation- CART
#install.packages("rpart")
#library(rpart)

# cart=rpart(poisox~ age+sex, data=df10 ,method='poisson')
# summary(cart)
# 
# df10_c=df10
# cart_ps=predict(cart)
# df10_c["propensity score"]=cart_ps
# df10_c %>% head()
# 
# df11_c=df11

df1_c=df1
df1_c["treat_f"]=as.factor(df1_c$treat)

cart2=rpart(treat_f~ age+sex, data=df1_c ,method='class')
summary(cart2)

cart_ps2=predict(cart2)
cart_ps2 %>% head()
df1_c %>% head()
df1_c$treat[95:100]

#add propensity score
for (i in 1:nrow(df1_c) ){

  if (df1_c$treat[i]==1){
    df1_c[i,"ps"]=cart_ps2[i,"1"]
  }
  else{
    df1_c[i,"ps"]=cart_ps2[i,"0"]
  }
}

#df1_c[65:70,]
```






```{r}
#add weighting- CART
zi=df1_c$treat
yi=df1_c$outcome
e=df1_c$ps

df1_c["ipw_wt"]= zi/e-(1-zi)/(1-e)
df1_c %>% head()

#ATE_ipw
ATE_ipw_cart=mean(zi*yi/e)-mean((1-zi)*yi/(1-e))
ATE_ipw_cart

#ATE_sipw
ATE_sipw_cart=sum(zi*yi/e)/sum(zi/e)-sum((1-zi)*yi/(1-e))/sum((1-zi)/(1-e))
ATE_sipw_cart
```

## Evaluate
```{r}

cov_balance_cart=data.frame(rep(0),row.names = "CART")

for(i in colnames(df1)){
  if(i!="treat" & i!="outcome"){
#    print(df1[i])
    t_weighted_mean=mean((df1[i]*df1_c$ipw_wt)[df1$treat==1,])
    c_weighted_mean=mean((df1[i]*df1_c$ipw_wt)[df1$treat==0,])
    weighted_mean_diff=abs(t_weighted_mean-c_weighted_mean)
    asam=weighted_mean_diff/sd((df1[i]*df1_c$ipw_wt)[df1$treat==1,])
    cov_balance_cart[i]=asam
  }
}
cov_balance_cart=cov_balance_cart[,-1]
cov_balance_cart$Average_ASAM=apply(cov_balance_cart,1,mean)
cov_balance_cart["Average_ASAM"]
```
```{r}
ATE_table= rbind(ATE_ipw_log,
      ATE_sipw_log,
      ATE_ipw_rf,
      ATE_sipw_rf,
      ATE_ipw_cart, 
      ATE_sipw_cart)

ATE_table
```

```{r, results='asis'}
writeLines("td, th { padding : 6px } th { background-color : brown ; color : white; border : 1px solid white; } td { color : brown ; border : 1px solid brown }", con = "mystyle.css")
knitr::kable(ATE_table, format = "html")
```


## Evaluation visualization
```{r}
ASAM_table=rbind(cov_balance_lo["Average_ASAM"] ,cov_balance_rf["Average_ASAM"] ,cov_balance_cart["Average_ASAM"])
ASAM_table

a=cbind(t(cov_balance_lo) ,t(cov_balance_rf) ,t(cov_balance_cart))
colnames(a)=c("LR","RF","CART")
boxplot(a, main="ASAM by propensity score estimation model",horizontal = TRUE)


b=cbind(df1_lo["ps"],df1_rf["ps"],df1_c["ps"])
colnames(b)=c("LR","RF","CART")
boxplot(b,horizontal = TRUE,main="Propensity score distribution by propensity score estimation model")

?par
par(mfcol=c(1,3))
title="Weight distribution for untreated observation by propensity score estimation model"
boxplot(df1_lo[(df1_lo$treat==0),"ipw_wt"],main="LR")
boxplot(df1_rf[(df1_lo$treat==0),"ipw_wt"],main="RF")
boxplot(df1_c[(df1_lo$treat==0),"ipw_wt"],main="CART")
```

## High blood pressure & Drinking

```{r}
df2$treat<-df2$treat-1
df2$sex<-df2$sex-1

table(df2$treat,df2$outcome)
# (df2$DC8_dg==0)
# sum(is.na(df2$treat))
# sum(is.na(df2$outcome))
# dim(df2)
# which(df2$treat==0) %>% head()
```

## ATE estimation - glm, random forest, cart

```{r}
#logsitic regression
df2_lo=df2

log_reg_ps=glm(treat~ age+sex, family = "binomial", data=df2)
logit_e_hat=predict(log_reg_ps)
print(logit_e_hat %>% head())

lo_ps=exp(logit_e_hat)/(1+exp(logit_e_hat))
print(lo_ps %>% head())


for (i in 1:nrow(df2_lo) ){
  if (df2_lo$treat[i]==1){
    df2_lo[i,"ps"]=lo_ps[i]
  }
  else{
    df2_lo[i,"ps"]=lo_ps[i]
  }
}
df2_lo %>% head()

#logistic regression - weighting

zi=df2_lo$treat
yi=df2_lo$outcome
e=df2_lo$ps

df2_lo["ipw_wt"]= zi/e-(1-zi)/(1-e)
df2_lo %>% head()

#ATE estimation
ATE_ipw_log=mean(zi*yi/e)-mean((1-zi)*yi/(1-e))
ATE_ipw_log

ATE_sipw_log=sum(zi*yi/e)/sum(zi/e)-sum((1-zi)*yi/(1-e))/sum((1-zi)/(1-e))
ATE_sipw_log

#Evaluate
cov_balance_lo=data.frame(rep(0),row.names = "logistic regression")

for(i in colnames(df2)){
  if(i!="treat" & i!="outcome"){
#    print(df2[i])
    t_weighted_mean=mean((df2[i]*df2_lo$ipw_wt)[df2$treat==1,])
    c_weighted_mean=mean((df2[i]*df2_lo$ipw_wt)[df2$treat==0,])
    weighted_mean_diff=abs(t_weighted_mean-c_weighted_mean)
    asam=weighted_mean_diff/sd((df2[i]*df2_lo$ipw_wt)[df2$treat==1,])
    cov_balance_lo[i]=asam
  }
}
cov_balance_lo=cov_balance_lo[,-1]
cov_balance_lo$Average_ASAM=apply(cov_balance_lo,1,mean)
cov_balance_lo["Average_ASAM"]

```

```{r}
# random forest - propensity score estimation
df2_rf=df2

rf=randomForest(treat~ age+sex, data=df2)
rf_ps=predict(rf)
print(rf_ps %>% head())


for (i in 1:nrow(df2_rf) ){
  if (df2_rf$treat[i]==1){
    df2_rf[i,"ps"]=rf_ps[i]
  }
  else{
    df2_rf[i,"ps"]=rf_ps[i]
  }
}
df2_rf %>% head()

#Random Forest- weighting

zi=df2_rf$treat
yi=df2_rf$outcome
e=df2_rf$ps

df2_rf["ipw_wt"]= zi/e-(1-zi)/(1-e)
df2_rf %>% head()

#ATE estimation
ATE_ipw_rf=mean(zi*yi/e)-mean((1-zi)*yi/(1-e))
ATE_ipw_rf

ATE_sipw_rf=sum(zi*yi/e)/sum(zi/e)-sum((1-zi)*yi/(1-e))/sum((1-zi)/(1-e))
ATE_sipw_rf

#Evaluate
cov_balance_rf=data.frame(rep(0),row.names = "Random Forest")

for(i in colnames(df2)){
  if(i!="treat" & i!="outcome"){
#    print(df2[i])
    t_weighted_mean=mean((df2[i]*df2_rf$ipw_wt)[df2$treat==1,])
    c_weighted_mean=mean((df2[i]*df2_rf$ipw_wt)[df2$treat==0,])
    weighted_mean_diff=abs(t_weighted_mean-c_weighted_mean)
    asam=weighted_mean_diff/sd((df2[i]*df2_rf$ipw_wt)[df2$treat==1,])
    cov_balance_rf[i]=asam
  }
}
cov_balance_rf=cov_balance_rf[,-1]
cov_balance_rf$Average_ASAM=apply(cov_balance_rf,1,mean)
cov_balance_rf["Average_ASAM"]

```

```{r}
#propensity score estimation- CART
#install.packages("rpart")
#library(rpart)

# cart=rpart(poisox~ age+sex, data=df20 ,method='poisson')
# summary(cart)
# 
# df20_c=df20
# cart_ps=predict(cart)
# df20_c["propensity score"]=cart_ps
# df20_c %>% head()
# 
# df21_c=df21

df2_c=df2
df2_c["treat_f"]=as.factor(df2_c$treat)

cart2=rpart(treat_f~ age+sex, data=df2_c ,method='class')
summary(cart2)

cart_ps2=predict(cart2)
cart_ps2 %>% head()
df2_c %>% head()
df2_c$treat[95:100]

#add propensity score
for (i in 1:nrow(df2_c) ){

  if (df2_c$treat[i]==1){
    df2_c[i,"ps"]=cart_ps2[i,"1"]
  }
  else{
    df2_c[i,"ps"]=cart_ps2[i,"0"]
  }
}

#df2_c[65:70,]
```



```{r}
#add weighting- CART
zi=df2_c$treat
yi=df2_c$outcome
e=df2_c$ps

df2_c["ipw_wt"]= zi/e-(1-zi)/(1-e)
df2_c %>% head()

#ATE_ipw
ATE_ipw_cart=mean(zi*yi/e)-mean((1-zi)*yi/(1-e))
ATE_ipw_cart

#ATE_sipw
ATE_sipw_cart=sum(zi*yi/e)/sum(zi/e)-sum((1-zi)*yi/(1-e))/sum((1-zi)/(1-e))
ATE_sipw_cart

#Evaluate
cov_balance_cart=data.frame(rep(0),row.names = "CART")

for(i in colnames(df2)){
  if(i!="treat" & i!="outcome"){
#    print(df2[i])
    t_weighted_mean=mean((df2[i]*df2_c$ipw_wt)[df2$treat==1,])
    c_weighted_mean=mean((df2[i]*df2_c$ipw_wt)[df2$treat==0,])
    weighted_mean_diff=abs(t_weighted_mean-c_weighted_mean)
    asam=weighted_mean_diff/sd((df2[i]*df2_c$ipw_wt)[df2$treat==1,])
    cov_balance_cart[i]=asam
  }
}
cov_balance_cart=cov_balance_cart[,-1]
cov_balance_cart$Average_ASAM=apply(cov_balance_cart,1,mean)
cov_balance_cart["Average_ASAM"]
```

```{r}
ATE_table= rbind(ATE_ipw_log,
      ATE_sipw_log,
      ATE_ipw_rf,
      ATE_sipw_rf,
      ATE_ipw_cart, 
      ATE_sipw_cart)

ATE_table
```

```
#Evaluate
```{r}
ASAM_table=rbind(cov_balance_lo["Average_ASAM"] ,cov_balance_rf["Average_ASAM"] ,cov_balance_cart["Average_ASAM"])
ASAM_table

a=cbind(t(cov_balance_lo) ,t(cov_balance_rf) ,t(cov_balance_cart))
boxplot(a, main="ASAM by propensity score estimation model")

```

### LaLonde dataset
```{r}
#
# haven::read_data("http://www.nber.org/~rdehejia/data/nsw_dw.dta")
#
# lalonde::nsw_dw
#
# data(LaLonde)
#
# print(data)
#
# y <- LaLonde$outcome
#
# trt <- LaLonde$treat
#
# x.varnames <- c("age", "educ", "black", "hisp", "white",
#                 "marr", "nodegr", "log.re75", "u75")
#
# # covariates
# data.x <- LaLonde[, x.varnames]
#
#
#
# # construct design matrix (with no intercept)
# x <- model.matrix(~ -1 + ., data = data.x)
#
# const.propens <- function(x, trt)
# {
#     mean.trt <- mean(trt == "Trt")
#     rep(mean.trt, length(trt))
# }
#
# subgrp_fit_w <- fit.subgroup(x = x, y = y, trt = trt,
#     loss = "logistic_loss_lasso",
#     propensity.func = const.propens,
#     cutpoint = 0,
#     type.measure = "auc",
#     nfolds = 10)
#
# summary(subgrp_fit_w)
```

